\thetitle

### 2008-06-26 paper.tex
New NMF intro paragraph.

We have addressed two out of three of the originally posed questions regarding traffic modeling; in this section, we address the third.
First, we have proposed a general approach for representing whole-network traffic without making assumptions of uniformity or marginality. To this end, we introduced the concept of a linear representation of traffic. The general matrix model entails a specific linear representation, such that network traffic patterns can easily be expressed as matrices. Our experimental results show that the \caps{GMM} is sufficiently realistic for generating wireless workloads derived from traces. When uniform or marginal assumptions are applied, however, sufficient realism eludes us, indicating that if we are to simplify and reduce the representation of network behaviors, we must do so by some other technique. In \Section{modeling-concepts}, we observed that application of uniformity and marginality assumptions to traffic instances is equivalent, in the \caps{GMM}, to right and left multiplication, respectively, by matrices of rank one. These transformations destroy almost all of the information in the original traffic matrix.
Conditional models are equivalent to left-multiplication by somewhat less degenerate matrices, but they achieve realism only by approaching the \caps{GMM} itself.

@@

It remains to show how we can simplify and reduce the representation of network behaviors without making unwarranted and manifestly false assumptions of uniformity of marginality of network behaviors. We present here preliminary results from applying nonnegative matrix factorization (\caps{NMF}) to the packet behaviors of real flows. Recall from \Section{general-matrix-model} that the matrices $\mat{Z}$ and $\mat{V}$ represent the packet size and inter-packet interval histograms of a collection of flows, with each row corresponding to a flow, and each column corresponding to a packet size or a range of inter-packet intervals. We apply \caps{NMF} to the concatenation of these matrices, $\mat{P}=[\mat{Z}~\mat{V}] \in \R^{f \times (d_z+d_v)}$. \caps{NMF} attempts to approximate $\mat{P}$ as the product of nonnegative matrices:
\begin{align}
\mat{P} \approx \mat{W}\mat{\Sigma}\mat{H},
\end{align}
where $\mat{W} \in \R^{f \times b}$ and $\mat{H} \in \R^{b \times (d_z+d_v)}$ are both row-stochastic, while
$\mat{\Sigma} \in \R^{b \times b}$ is diagonal with descending diagonal entries.
$b\in\N$ is the rank of this factorization, which must be determined somehow as well.
What is the interpretation of such a factorization?
The rows of $\mat{H}$ are \emph{basic behaviors}---pairs of packet size and inter-packet interval distributions, such that each flow's packet behavior can be approximated as a linear combination of these basic behaviors.
Specifically, the matrix $\mat{M}\mat{\Sigma}$ gives the mixing coefficients for each flow.
The diagonal entries of $\mat{\Sigma}$ have a special meaning as well:
they are the number of packets associated with each basic behavior.
Thus the basic behaviors are sorted in descending order of how much traffic they explain.
% Note that the factorization does not associate basic behaviors with individual packets;
% rather, total flow behavior is explained as a mixture of basic behaviors, so the fraction of packets in each flow explained by a given behavior is known, while the specific packets are unknown.

>>>>

% We have addressed two out of three of the originally posed questions regarding traffic modeling; in this section, we address the third.
% First, we have proposed a general approach for representing whole-network traffic without making assumptions of uniformity or marginality. To this end, we introduced the concept of a linear representation of traffic. The general matrix model entails a specific linear representation, such that network traffic patterns can easily be expressed as matrices. Our experimental results show that the \caps{GMM} is sufficiently realistic for generating wireless workloads derived from traces. When uniform or marginal assumptions are applied, however, sufficient realism eludes us, indicating that if we are to simplify and reduce the representation of network behaviors, we must do so by some other technique. In \Section{modeling-concepts}, we observed that application of uniformity and marginality assumptions to traffic instances is equivalent, in the \caps{GMM}, to right and left multiplication, respectively, by matrices of rank one. These transformations destroy almost all of the information in the original traffic matrix.
% Conditional models are equivalent to left-multiplication by somewhat less degenerate matrices, but they achieve realism only by approaching the \caps{GMM} itself.

@@

It remains to show how we can simplify and reduce the representation of network behaviors without making unwarranted and manifestly false assumptions of uniformity of marginality of network behaviors.
In \Section{modeling-concepts}, we observed that application of uniformity and marginality assumptions to traffic instances is equivalent, in the \caps{GMM}, to right and left multiplication, respectively, by matrices of rank one. These transformations destroy almost all of the information in the original traffic matrix.
Conditional models are equivalent to left-multiplication by somewhat less degenerate matrices, but they achieve realism only by approaching the \caps{GMM} itself.
In this section we describe matrix factorization, which provides an approximate low-rank representation of the original traffic.
This technique not only allows us to model and analyze traffic without destroying the information content, but also leads us to a deep analytical understanding of network behaviors that matches our intuitive understanding of typical modes of network behavior.

We present here preliminary results from applying nonnegative matrix factorization (\caps{NMF}) to the packet behaviors of real flows. Recall from \Section{general-matrix-model} that the matrices $\mat{Z}$ and $\mat{V}$ represent the packet size and inter-packet interval histograms of a collection of flows, with each row corresponding to a flow, and each column corresponding to a packet size or a range of inter-packet intervals. We apply \caps{NMF} to the concatenation of these matrices, $\mat{P}=[\mat{Z}~\mat{V}] \in \R^{f \times (d_z+d_v)}$. \caps{NMF} attempts to approximate $\mat{P}$ as the product of nonnegative matrices:
\begin{align}
\mat{P} \approx \mat{W}\mat{\Sigma}\mat{H},
\end{align}
where $\mat{W} \in \R^{f \times b}$ and $\mat{H} \in \R^{b \times (d_z+d_v)}$ are both row-stochastic, while
$\mat{\Sigma} \in \R^{b \times b}$ is diagonal with descending diagonal entries.
$b\in\N$ is the rank of this factorization, which must be determined somehow as well.
What is the interpretation of such a factorization?
The rows of $\mat{H}$ are \emph{basic behaviors}---pairs of packet size and inter-packet interval distributions, such that each flow's packet behavior can be approximated as a linear combination of these basic behaviors.
Specifically, the matrix $\mat{M}\mat{\Sigma}$ gives the mixing coefficients for each flow.
The diagonal entries of $\mat{\Sigma}$ have a special meaning as well:
they are the number of packets associated with each basic behavior.
Thus the basic behaviors are sorted in descending order of how much traffic they explain.
% Note that the factorization does not associate basic behaviors with individual packets;
% rather, total flow behavior is explained as a mixture of basic behaviors, so the fraction of packets in each flow explained by a given behavior is known, while the specific packets are unknown.
